Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2019-06-07 GWLDCI.tex
ignore:testing
edits in section 4 of paper 1

\begin{dfn}
  Let $x_1, \ldots, x_n$ define a convex $n$-gon.  Let $T$ be the set of triangulations of this $n$-gon.  For each $t\in T$ define $s_t$ to be the point in $\mathbb{R}^n$ whose $i$th coordinate is the sum of the areas of all triangles of $t$ which are incident to $x_i$.  Let $A_n$ be the convex hull of the $s_t$.
\end{dfn}
This is the construction of the secondary polytope (so $A_n$ is the secondary polytope of $x_1, \ldots, x_n$), see \cite{Ziegler} Definition 9.9.  In general the $x_1, \ldots, x_n$ may be any points in any $\mathbb{R}^d$.  Triangulations in this more general context are decompositions into simplices and the area of the triangles is replaced by the volume of the simplices.

@@

  Consider now an edge whose quadrilateral has corners $x_i$, $x_j$, $x_k$, and $x_l$ in that order.  Let $t_1$ and $t_2$ be the triangulations corresponding to the two ends of the edge where $t_1$ has a diagonal from $x_j$ to $x_l$ and $t_2$ has a diagonal from $x_i$ to $x_k$.  From the definition of $A_n$ we see that $s_{t_1}$ and $s_{t_2}$ only differ at coordinates $i$, $j$, $k$, and $l$, and all of the triangles incident to any of these vertices other than the one coming from triangulating $x_i, x_j, x_k, x_l$ are the same in both $t_1$ and $t_2$.  Let $a$ be the area of the quadrilateral and let $a(i,j,k)$ be the area of the triangle with corners $x_i$, $x_j$ and $x_k$, and similarly for other indices then $s_{t_1}$ on coordinates $i$, $j$, $k$, $l$ is $(a(i,j,l) + b_i, a+b_j, a(k,l,j)+b_k, a+b_l)$ and $s_{t_2}$ on these coordinates is $(a + b_i, a(j,k,i)+b_j, a+b_k, a(l,i,k)+b_l$, where the $b_i$ are the sums of the areas of the relevant triangles which are common to both triangulations.  Thus a vector parallel to this edge is the vector which is $(a(i,j,l) - a, a-a(j,k,i), a(k,l,j)-a, a-a(l,i,k)) = (-a(j,k,l), a(i,k,l), -a(i,j,l), a(i,j,k))$ on $i$, $j$, $k$, $l$ and $0$ in all other coordinates.  This is illustrated in Figure~\ref{triangulate}.

@@

  Returning to $\tau(W)$ and $\tau(X)$, since the direction vector of an edge in $A_n$ depends only on its quadrilateral and the bijection between the edges bounding $\tau(W)$ and the edges bounding $\tau(X)$ preserves the quadrilateral, we have the same direction vectors bounding $\tau(W)$ as $\tau(X)$.  Furthermore, the bijection also maintains the relative configuration of the edges and so $\tau(W)$ and $\tau(X)$ are parallel.
\end{proof}

The following lemma is useful for proving the converse of the above. We also need the following notation.  Given a polygon dissection $D$ let $V_{>3}(D)$ be the set of vertices of the polygon which are on at least one non-triangle face of $D$.  Note that $D$ may have many non-triangle faces and so $V_{>3}(D)$ will be the union of the vertices of each non-triangle face.

\begin{lem}\label{lem good quads}
  Let $D_1$ and $D_2$ be two dissections of the convex polygon defined by the vertices $x_1, \ldots, x_n$.  Suppose $V_{>3}(D_1) = V_{>3}(D_2)$, but that the $D_1$ and $D_2$ do not have the same set of non-triangle faces.  Then there are four vertices $x_i$, $x_j$, $x_k$, $x_l$ such that in one of $D_1$ or $D_2$ all four vertices are on the same non-triangle face while in the other there is a diagonal of the dissection such that there is one side of the diagonal with exactly one of $\{x_i, x_j, x_k, x_l\}$ (so the others are end points of this diagonal or strictly on the other side).
\end{lem}

@@

Assume for a contradiction that no quadrilateral as in the statement exists, and let $d$ and $f$ be as above.  Then $f$ must have two vertices strictly on each side of $d$, say $x_i$ and $x_j$ on one side and $x_k$, $x_l$ on the other side (and in that order).  Consider the face incident to $d$ on the $x_i, x_j$ side of $d$.  Suppose this face is a triangle, then its other vertex cannot be in $[x_i, x_j]$ as otherwise taking one of the other edges of the triangles would give a configuration as in the statement.  Suppose now that this face is a quadrilateral, again no other vertex of this face can be in $(x_i, x_j)$ as otherwise again taking one of the other edges of the face as $d$ we'd get a quadrilateral as in the statement, the same holds if one of the other vertices of the face is $x_i$ or $x_j$ but the other is not.  If the two other vertices of the face are $x_i$ and $x_j$, then we also get a configuration as in the statement but now with the four vertices being the two ends of $q$ and $x_i, x_j$ and with one of the edges of $f$ as the diagonal.  Thus we can conclude that none of the vertices of this face are in $[x_i, x_j]$.  Also if the face has three vertices on one side of $f$ and one on the other then we get a quadrilateral as in the statement again with this face giving the four vertices and the diagonal is a side of $f$.  In this case the face is either a triangle with its third vertex of the triangle must be in $(x_j,x_k)$ or in $(x_l,x_i)$ or a quadrilateral with one more vertex in $(x_j, x_k)$ and one more vertex in $(x_l, x_i)$.  The three possibilities are illustrated in Figure~\ref{inductive_step}. Taking all this together, whether the face is a triangle or a quadrilateral, replacing $d$ with one of the other edges of the face we get a new diagonal $d'$ which also separates $f$ into $x_i, x_j$ and $x_k, x_l$ but is strictly closer to $[x_i, x_j]$ than $d$ was.  The same argument holds for $d'$, but the polygon is finite, so this process cannot go on indefinitely, contradicting that no quadtilateral as in the statement exists.
\end{proof}

>>>>

\begin{dfn} \label{Andfn}
  Let $x_1, \ldots, x_n$ define a convex $n$-gon.  Let $T$ be the set of triangulations of this $n$-gon.  For each $t\in T$ define $s_t$ to be the point in $\mathbb{R}^n$ whose $i$th coordinate is the sum of the areas of all triangles of $t$ which are incident to $x_i$.  Let $A_n$ be the convex hull of the $s_t$.
\end{dfn}
This is the construction of the secondary polytope (so $A_n$ is the secondary polytope of $x_1, \ldots, x_n$), see \cite{Ziegler} Definition 9.9.  In general the $x_1, \ldots, x_n$ may be any points in any $\mathbb{R}^d$.  Triangulations in this more general context are decompositions into simplices and the area of the triangles is replaced by the volume of the simplices.

@@

Consider now an edge in $A_n$ whose quadrilateral is defined by the vertices of $\tau(W)$ or $\tau(X), $x_i$, $x_j$, $x_k$, and $x_l$ in that order.  Let $t_1$ and $t_2$ be the triangulations corresponding to the two ends of the edge where $t_1$ has a diagonal from $x_j$ to $x_l$ and $t_2$ has a diagonal from $x_i$ to $x_k$.  From Definition \ref{Andfn}, defining $A_n$,  we see that $s_{t_1}$ and $s_{t_2}$ only differ at coordinates $i$, $j$, $k$, and $l$. All of the triangles incident to any of these vertices other than the one coming from triangulating $x_i, x_j, x_k, x_l$ are the same in both $t_1$ and $t_2$.  Let $a$ be the area of the quadrilateral and let $a(i,j,k)$ be the area of the triangle with corners $x_i$, $x_j$ and $x_k$, and similarly for other indices then $s_{t_1}$ on coordinates $i$, $j$, $k$, $l$ is $(a(i,j,l) + b_i, a+b_j, a(k,l,j)+b_k, a+b_l)$ and $s_{t_2}$ on these coordinates is $(a + b_i, a(j,k,i)+b_j, a+b_k, a(l,i,k)+b_l$, where the $b_i$ are the sums of the areas of the relevant triangles which are common to both triangulations.  Thus a vector parallel to this edge is the vector which is $(a(i,j,l) - a, a-a(j,k,i), a(k,l,j)-a, a-a(l,i,k)) = (-a(j,k,l), a(i,k,l), -a(i,j,l), a(i,j,k))$ on $i$, $j$, $k$, $l$ and $0$ in all other coordinates.  This is illustrated in Figure~\ref{triangulate}.

@@

  Returning to $\tau(W)$ and $\tau(X)$, since the direction vector of an edge in $A_n$ depends only on its quadrilateral and the bijection between the edges bounding $\tau(W)$ and the edges bounding $\tau(X)$ preserves the quadrilateral, we have the same direction vectors bounding $\tau(W)$ as $\tau(X)$.  Furthermore, the bijection also maintains the relative configuration of the edges and so the edges corresponding to $\tau(W)$ and $\tau(X)$ are parallel.
\end{proof}

The following lemma is useful for proving the converse of the above. We also need the following notation.  Given a polygon dissection $D$ of an $n$-gon, let $V_{>3}(D)$ be the set of vertices of the polygon which are on at least one non-triangle face of $D$.  Note that $D$ may have many non-triangle faces and so $V_{>3}(D)$ will be the union of the vertices of each non-triangle face.

\begin{lem}\label{lem good quads}
Let $D_1$ and $D_2$ be two dissections of a convex $n$-gon defined by the vertices $x_1, \ldots, x_n$.  Suppose $V_{>3}(D_1) = V_{>3}(D_2)$, but that the $D_1$ and $D_2$ do not have the same set of non-triangle faces.  Then there are four vertices $x_i$, $x_j$, $x_k$, $x_l$ satisfying the following conditions:\begin{enumerate} \item in either $D_1$ or $D_2$ all four vertices are on the same non-triangle face \item in the other disection, there is a diagonal of the dissection such that exactly one of $\{x_i, x_j, x_k, x_l\}$ lies on one side (so the others are either end points of this diagonal or strictly on the other side). \end{enumerate}
\end{lem}

@@

This proof is by a symmetrical consruction, where we give a proof with respect to a single edge in $f$, leaving the rest of the argument as read due to the symmetry.

Assume for a contradiction that no quadrilateral as in the statement exists, and let $d$ and $f$ be as above.  Then $f$ must have two vertices strictly on each side of $d$, say $x_i$ and $x_j$ on one side and $x_k$, $x_l$ on the other side (and in that order).  Consider the face in $D_1$ incident to $d$ on the $x_i, x_j$ side of $d$. Call this $g$. Suppose this face is a triangle, then its other vertex cannot be in \st{$[x_i, x_j]$} the cyclic interval bounded by $x_i$ and $x_j$ (i.e. $[x_i, x_j]$) as in this case taking one of the other edges of the triangle $g$ would give a configuration as in the statement.  Suppose then that this face, $g$, is a quadrilateral. Again no other vertex of this face can be in $(x_i, x_j)$ as otherwise again replacing $d$ with one of the other edges of $g$ gives a quadrilateral as in the statement. The same holds if one of the other vertices of $g$ is $x_i$ or $x_j$ but the other is not.  If both the other vertices of $g$ are $x_i$ and $x_j$, then we also get a configuration as in the statement but now with the four vertices being those defining $g$ and the diagonal is one of the edges of $f$. (For instance $d$ can now be the edge connecting $x_l$ to $x_i$ or $x_j$ to $x_k$. Thus we can conclude that none of the vertices of this face are in $[x_i, x_j]$.

Suppose the face $g$ has one vertex in $(x_j, ,x_k)$ and the rest in $(x_l, x_i)$ (or vice versa). If $g$ is a quadrilateral, its four vertices satisfy the condition with the edge between $x_k$ and $x_j$ in $f$ serving as an example we get a quadrilateral as in the statement again with this face giving the four vertices and the diagonal is a side of $f$.  If $g$ a triangle, let $d'$ be the other diagonal across $f$. These are the first two cases illustrated in Figure~\ref{inductive_step}.

Finally, if the face $g$ has two vertices in $(x_j, x_k)$ and two in $(x_l, x_i)$, let $d'$ be the the other edge of $g$ that forms a diagonal across $f$. In these last two cases (those represented in Figure~\ref{inductive_step}), replace $d$ with $d'$, which also separates $f$ into $x_i, x_j$ and $x_k, x_l$ \hlfix{but is strictly closer to $[x_i, x_j]$ than $d$ was}{why is this the case?}.  The same argument holds for $d'$, but the polygon is finite, so this process cannot go on indefinitely, contradicting that no quadtilateral as in the statement exists.
\end{proof}
